Beyond the Iceberg Hypothesis: \\Opening the Black Box of Transport Costs

### 2016-09-23 redaction/Presentations/SeminaireLEDa_Sept2016/slides_LEDa.tex
Changements slides LEDa + pgm stats des

\begin{frame}[label=slide_results_summary]
\frametitle{Result 1: Estimating transport costs over time}
For both the ad-valorem and the additive components \vspace{0.1cm}
\begin{itemize}
\item[-] \footnotesize{Average values over 1974-2013, in percent of the export price \hyperlink{app_results_summary}{\beamergotobutton{More}}  } \vspace{0.1cm}
\end{itemize}
\begin{table}[htbp]
  \centering
  \scriptsize{
  %\caption{Transport costs estimates: Summary \label{tab:summary_results}}
  \begin{center}
    \begin{tabular}{l|cc|cc}
      \hline \hline
    %\multicolumn{5}{c}{Mean value over 1974-2013}   \\
    \# digit & \multicolumn{2}{c}{3 digits} & \multicolumn{2}{c}{4 digits} \\ \hline
    Mode  & Vessel & Air & Vessel & Air \\ \hline
    \multicolumn{5}{l}{With only Ad-Valorem Costs ($\widehat{\tau}^{ice}$)}  \\ \hline
    Mean  & \bf{5.8} & \bf{5.1} & 6.0 & 4.9 \\
    Median & 5.1 & 4.2 & 5.2 & 3.7 \\ \hline
    %Std   & 0.032 & 0.042 & 0.036 & 0.045 \\
    %Min. value & 1.003 & 1.001 & 1.003 & 1.000 \\
    %Max. value & 1.304 & 1.685 & 1.408 & 2.051 \\ \hline
    \multicolumn{5}{l}{With Additive \& Ad-Valorem Costs } \\ \hline
   \multicolumn{5}{l}{\textit{Ad-valorem term} ($\widehat{\tau}^{adv}$)}\\ \hline
    Mean  & \bf{3.2} & \bf{2.5} & 3.3 & 2.4 \\
    Median & 2.8 & 1.8 & 2.8 & 1.6 \\ \hline
    %Std   & 0.023 & 0.023 & 0.025 & 0.026 \\
    %Min. value & 1.001 & 1.000 & 1.000 & 1.000 \\
    %Max. value & 1.227 & 1.474 & 1.264 & 1.537 \\ \hline
    \multicolumn{5}{l}{\textit{Additive term }($\widehat{t}^{add}/\widetilde{p}$)} \\ \hline
    Mean  & \bf{2.9} & \bf{1.8} & 2.8 & 1.9 \\
    Median & 1.9 & 0.7 & 1.7 & 0.8 \\ \hline
    %Std   & 0.041 & 0.034 & 0.039 & 0.034 \\
    %Min. value & 0.000 & 0.000 & 0.000 & 0.000 \\
    %Max. value & 2.941 & 13.303 & 3.197 & 11.440 \\ \hline
    %   \multicolumn{5}{l}{\textit{Data} }\\ \hline
    %Mean &5.6 &4.0 &5.7 & 4.0\\
    %Median  &4.4 & 1.9 &4.5 & 2.0 \\  \hline
    \# obs. & 29279 & 28207 & 29317 & 27680 \\
    \# origin country & 188 & 191 & 188 & 189 \\
    \# products & 230 & 211 & 666 & 567 \\    \hline \hline
  \end{tabular}
%\parbox[l]{8cm}{\tiny{Notes: Statistics are obtained weighting each observation by its value.}}
%. The additive term is expressed in fraction of fab price. ($^\ast$): Four 4-digit estimation: 0n selected years. ($^{\ast \ast}$): 1989 omitted in 3 digit estimation for air.}}
\end{center}}
  \end{table}%
 \end{frame}

@@

\begin{frame}
\frametitle{Excluding the composition effects}
\begin{itemize}
\item For the ad-valorem component, we estimate the following equation:
\footnotesize
\begin{eqnarray}
%\widehat{\tau}_{ikt}&=&\delta\times \exp\left(\sum_{i \neq \text{AFG}}\alpha_i.\mathbb{1}_i\right).\exp\left(\sum_{k\neq \text{011}}\beta_k.\mathbb{1}_k\right).\exp\left(\sum_{t \neq 1974}\gamma_t.\mathbb{1}_t\right) .\exp\left(\epsilon_{ikt}\right) \notag \\
\ln(\widehat{\tau}_{ikt})&=&\delta +\underbrace{\sum_{i \neq \text{AFG}}\alpha_i.\mathbb{1}_i }_{\ln(\widetilde{\tau}_i)}+ \underbrace{\sum_{k\neq \text{011}}\beta_k.\mathbb{1}_k }_{\ln(\widetilde{\tau}_k)}+ \underbrace{\sum_{t \neq 1974}\gamma_t.\mathbb{1}_t}_{\text{``Pure'' TC}}+\epsilon_{ikt} \label{eq:compeffects_mult}
\end{eqnarray}
\normalsize
\begin{itemize}
\item[-] With $\widehat{\tau}_{ikt} = \widehat{\tau}^{ice}_{ikt}, \widehat{\tau}^{adv}_{ikt}$ previously obtained  \vspace{0.1cm}
\end{itemize}
\item For the additive component:
\footnotesize
\begin{eqnarray}
%\widehat{t}_{ikt}&=&\left( \prod_{i \neq \text{ARG}}  \alpha_i.\mathbb{1}_i+\prod_{k}\beta_k.\mathbb{1}_k\right).\exp\left(\sum_{t \neq 1974}\gamma_t.\mathbb{1}_t\right) .\exp\left(\epsilon_{ikt}\right) \notag \\
\ln(\widehat{t}_{ijt})&=&\ln\left( \underbrace{\prod_{i \neq \text{ARG}}  \alpha_i.\mathbb{1}_i}_{\widetilde{t}_i}+ \underbrace{\prod_{k}\beta_k.\mathbb{1}_k}_{\widetilde{t}_k}\right) + \underbrace{\sum_{t \neq 1974}\gamma_t.\mathbb{1}_t}_{\text{``Pure'' TC}}+\epsilon_{ijt} \label{eq:compeffects_add}
\end{eqnarray}
\normalsize
\item Underlying rationale
\footnotesize
\begin{itemize}
\item[-] Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}): Preserve our specification of the ad-valorem and the additive costs (Equation (\ref{eq:specifTC})) \vspace{0.1cm}
\item[-] Equation (\ref{eq:compeffects_mult}) estimated using OLS, Equation (\ref{eq:compeffects_add}) using non-linear least squares (by transport mode)
    \normalsize
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\begin{itemize}
\item Exclude the composition effects of transport costs changes  \vspace{0.1cm}
\item[$\Leftrightarrow$] Isolate the change in the time dimension  \vspace{0.1cm}
\begin{itemize}
\item[-] From the ad-valorem component estimation (Equation (\ref{eq:compeffects_mult})), build the variable $\Gamma_t$ ($\forall~t > 1974$):
\begin{eqnarray*}
&&\Gamma_t = 100\frac {\bar{\tau}_{t}-1} {\bar{\tau}_{1974}-1} \\
\Leftrightarrow && \Gamma_t= \frac {\bar{\tau}_{1974}.\exp(\gamma_t)-1} {\bar{\tau}_{1974}-1}
\end{eqnarray*}
\begin{itemize}
\item[$\ast$] with $\bar{\tau}_{1974} = \exp(\delta+\sum_i \alpha_i + \sum_k \beta_k$) the mean TC in 1974
\end{itemize}
\item[-] For the additive cost, we build the variable ($\forall~t > 1974$)
$$\Gamma^{add}_t = 100 \exp(\gamma_t)$$
\end{itemize}
\item The $\Gamma^{add}_t$ and $\Gamma_t$ series: Interpretation in percentage changes
\begin{itemize}
\item[] with an initial value of 100 for $t=1974$
\end{itemize}
\end{itemize}

@@

>>>>

\begin{frame}[label=slide_results_summary]
\frametitle{Result 1: Estimating transport costs over time}
For both the ad-valorem and the additive components \vspace{0.1cm}
\begin{itemize}
\item[-] \footnotesize{Average values over 1974-2013, in percent of the export price \hyperlink{app_results_summary}{\beamergotobutton{More}}  } \vspace{0.1cm}
\end{itemize}
\begin{table}[htbp]
  \centering
  \scriptsize{
  %\caption{Transport costs estimates: Summary \label{tab:summary_results}}
  \begin{center}
    \begin{tabular}{l|cc|cc}
      \hline \hline
    %\multicolumn{5}{c}{Mean value over 1974-2013}   \\
    \# digit & \multicolumn{2}{c}{3 digits} & \multicolumn{2}{c}{4 digits} \\ \hline
    Mode  & Vessel & Air & Vessel & Air \\ \hline
    \multicolumn{5}{l}{With only Ad-Valorem Costs ($\widehat{\tau}^{ice}$)}  \\ \hline
    Mean  & \bf{5.8} & \bf{5.1} & 6.0 & 4.9 \\
    Median & 5.1 & 4.2 & 5.2 & 3.7 \\ \hline
    %Std   & 0.032 & 0.042 & 0.036 & 0.045 \\
    %Min. value & 1.003 & 1.001 & 1.003 & 1.000 \\
    %Max. value & 1.304 & 1.685 & 1.408 & 2.051 \\ \hline
    \multicolumn{5}{l}{With Additive \& Ad-Valorem Costs } \\ \hline
   \multicolumn{5}{l}{\textit{Ad-valorem term} ($\widehat{\tau}^{adv}$)}\\ \hline
    Mean  & \bf{3.2} & \bf{2.5} & 3.3 & 2.4 \\
    Median & 2.8 & 1.8 & 2.8 & 1.6 \\ \hline
    %Std   & 0.023 & 0.023 & 0.025 & 0.026 \\
    %Min. value & 1.001 & 1.000 & 1.000 & 1.000 \\
    %Max. value & 1.227 & 1.474 & 1.264 & 1.537 \\ \hline
    \multicolumn{5}{l}{\textit{Additive term }($\widehat{t}^{add}/\widetilde{p}$)} \\ \hline
    Mean  & \bf{2.9} & \bf{1.8} & 2.8 & 1.9 \\
    Median & 1.9 & 0.7 & 1.7 & 0.8 \\ \hline
    %Std   & 0.041 & 0.034 & 0.039 & 0.034 \\
    %Min. value & 0.000 & 0.000 & 0.000 & 0.000 \\
    %Max. value & 2.941 & 13.303 & 3.197 & 11.440 \\ \hline
       \multicolumn{5}{l}{\textit{Data} }\\ \hline
    Mean &5.3 &5.0 & 5.6&3.9 \\
    Median  &4.3 & 2.0 &4.4 & 1.9 \\  \hline
    \# obs. & 29279 & 28207 & 29317 & 27680 \\
    \# origin country & 188 & 191 & 188 & 189 \\
    \# products & 230 & 211 & 666 & 567 \\    \hline \hline
  \end{tabular}
%\parbox[l]{8cm}{\tiny{Notes: Statistics are obtained weighting each observation by its value.}}
%. The additive term is expressed in fraction of fab price. ($^\ast$): Four 4-digit estimation: 0n selected years. ($^{\ast \ast}$): 1989 omitted in 3 digit estimation for air.}}
\end{center}}
  \end{table}%
 \end{frame}

@@

\begin{frame}
\frametitle{Excluding the composition effects}
\begin{itemize}
\item For the ad-valorem component, we estimate the following equation:
\footnotesize
\begin{eqnarray}
%\widehat{\tau}_{ikt}&=&\delta\times \exp\left(\sum_{i \neq \text{AFG}}\alpha_i.\mathbb{1}_i\right).\exp\left(\sum_{k\neq \text{011}}\beta_k.\mathbb{1}_k\right).\exp\left(\sum_{t \neq 1974}\gamma_t.\mathbb{1}_t\right) .\exp\left(\epsilon_{ikt}\right) \notag \\
\ln(\widehat{\tau}_{ikt})&=&\delta +\underbrace{\sum_{i \neq \text{AFG}}\alpha_i.\mathbb{1}_i }_{\ln(\widetilde{\tau}_i)}+ \underbrace{\sum_{k\neq \text{011}}\beta_k.\mathbb{1}_k }_{\ln(\widetilde{\tau}_k)}+ \underbrace{\sum_{t \neq 1974}\gamma_t.\mathbb{1}_t}_{\text{Time trend}}+\epsilon_{ikt} \label{eq:compeffects_mult}
\end{eqnarray}
\normalsize
\begin{itemize}
\item[-] With $\widehat{\tau}_{ikt} = \widehat{\tau}^{ice}_{ikt}, \widehat{\tau}^{adv}_{ikt}$ previously obtained  \vspace{0.1cm}
\end{itemize}
\item For the additive component:
\footnotesize
\begin{eqnarray}
%\widehat{t}_{ikt}&=&\left( \prod_{i \neq \text{ARG}}  \alpha_i.\mathbb{1}_i+\prod_{k}\beta_k.\mathbb{1}_k\right).\exp\left(\sum_{t \neq 1974}\gamma_t.\mathbb{1}_t\right) .\exp\left(\epsilon_{ikt}\right) \notag \\
\ln(\widehat{t}_{ijt})&=&\ln\left(\delta +  \underbrace{\sum_{i \neq \text{ARG}}  \alpha_i.\mathbb{1}_i}_{\widetilde{t}_i}+ \underbrace{\sum_{k \neq \text{011}}\beta_k.\mathbb{1}_k}_{\widetilde{t}_k}\right) + \underbrace{\sum_{t \neq 1974}\gamma_t.\mathbb{1}_t}_{\text{Time trend}}+\epsilon_{ijt} \label{eq:compeffects_add}
\end{eqnarray}
\normalsize
\item Underlying rationale 
\footnotesize
\begin{itemize}
\item[-] Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}): Preserve our specification of the ad-valorem and the additive costs (Equation (\ref{eq:specifTC})) \vspace{0.1cm}
\item[-] Equation (\ref{eq:compeffects_mult}) estimated using OLS, Equation (\ref{eq:compeffects_add}) using non-linear least squares (by transport mode)
    \normalsize
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}[label = slide_compeffects]
\begin{itemize}
\item Exclude the composition effects of transport costs changes  \hyperlink{app_compeffects}{\beamergotobutton{More details}}\vspace{0.1cm}
\item[$\Leftrightarrow$] Isolate the change in the time dimension  \vspace{0.1cm}
\begin{itemize}
\item[-] From the ad-valorem component estimation (Equation (\ref{eq:compeffects_mult})), build the variable $\Gamma_t$ ($\forall~t > 1974$):
\begin{equation*}
 \Gamma_t= \frac {\bar{\tau}_{1974}.\exp(\gamma_t)-1} {\bar{\tau}_{1974}-1}
\end{equation*}
\begin{itemize}
\item[$\ast$] with $\bar{\tau}_{1974} = \exp(\delta+\sum_i \alpha_i + \sum_k \beta_k$) the mean TC in 1974
\end{itemize}
\item[-] For the additive cost, we build the variable ($\forall~t > 1974$)
$$\Gamma^{add}_t = 100 \exp(\gamma_t)$$
\end{itemize}
\item The $\Gamma^{add}_t$ and $\Gamma_t$ series: Interpretation in percentage changes
\begin{itemize}
\item[] with an initial value of 100 for $t=1974$
\end{itemize}
\end{itemize}

@@

\begin{frame}[label=app_compeffects]
\frametitle{Excluding the composition effects: More details}
\begin{itemize}
\item \textbf{For the ad-valorem cost}, rewriting Equation (\ref{eq:compeffects_mult}) :
\footnotesize
\begin{equation*}
\widehat{\tau}_{ikt}=\exp\left(\delta + \sum_{i \neq \text{AFG}}\alpha_i.\mathbb{1}_i+\sum_{k\neq \text{011}}\beta_k.\mathbb{1}_k\right).\exp\left(\sum_{t \neq 1974}\gamma_t.\mathbb{1}_t\right) .\exp\left(\epsilon_{ikt}\right)
\end{equation*}
\normalsize
\begin{itemize}
\item[-] From which we deduce after estimation: $\widehat{\tau}_{ik74} = \exp(\delta +\alpha_i +\beta_k)$ \vspace{0.1cm}
\item[-] And, for any year $t > 1974$: $\widehat{\tau}_{ikt} = \exp(\delta +\alpha_i +\beta_k)\times \exp(\gamma_t)$ \vspace{0.1cm}
\item[-] Which implies:
\begin{equation*}
\widehat{\tau}_{ikt} = \widehat{\tau}_{ik74}\times \exp(\gamma_t)
\end{equation*}
\end{itemize}
\item With $\tau>1$, rewrite things to get the percentage change between years 1974 and $t$:
\footnotesize
$$\Gamma_{ikt} = 100.\frac{\widehat{\tau}_{ikt}-1}{\widehat{\tau}_{ik74}-1} = 100.\frac{\widehat{\tau}_{ik74}\exp(\gamma_t)-1}{\widehat{\tau}_{ik74}-1}$$
\normalsize
\begin{itemize}
\item[-] With $\Gamma_{ikt}$ an index of transport costs in $t$ (relative to 1974) \vspace{0.1cm}
\item[-] Only depend on the cost observed in 1974 and the time trend \vspace{0.1cm}
\item[-] But, specific to a product-origin country pair  \vspace{0.1cm}

\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\begin{itemize}
\item Choose to pick as reference the mean value of the cost in 1974
\begin{itemize}
\item[-] Build the index $\Gamma_t$ such that:
\footnotesize
\begin{equation*}
 \Gamma_t= 100.\frac {\bar{\tau}_{1974}.\exp(\gamma_t)-1} {\bar{\tau}_{1974}-1}
\end{equation*}
\item[-] With $\bar{\tau}_{1974} = \exp(\delta + \sum_i \alpha_i + \sum_k \beta_k$) the mean TC in 1974 \vspace{0.2cm}
\end{itemize}
\item \textbf{For the additive component}, after estimating Equation (\ref{eq:compeffects_add}):
\footnotesize
\begin{eqnarray*}
\widehat{t}_{ik74}&= & \delta + \alpha_i+ \beta_k \\
\widehat{t}_{ikt}&=&\left(\delta + \alpha_i+ \beta_k\right).\exp(\gamma_t),\qquad \forall t > 1974
\end{eqnarray*}
\normalsize

\begin{itemize}
\item[-] From which we deduce: 
$$\widehat{t}_{ikt} = \widehat{t}_{ik74} \times \exp(\gamma_t)$$
\item[-] With $t >0$, obtain the percentage change from 1974 through:
\footnotesize
\begin{equation*}
\Gamma^{add}_{ikt} = 100\frac{\widehat{t}_{ikt}}{\widehat{t}_{ik74}} = 100\exp(\gamma_t)
\end{equation*}
\normalsize
\item Independent of the $i,k$ dimension, such that we can rewrite:
$$\Gamma^{add}  = 100\exp(\gamma_t) $$
\end{itemize}
\end{itemize}
\hyperlink{slide_compeffects}{\beamergotobutton{Back to slide}}
\end{frame}
